---
title: "Visualization"
output: .
  revealjs::revealjs_presentation:
    theme: moon
    highlight: pygments
    center: false
    transition: fade
    smart: true
    background_transition: slide
    css: workshop.css
---


## 

<div style='text-align: center; font-size:32'><h1>Interactive and Visual Data Exploration<h1></div>


## The Viewer

In addition to the "Plots" pane, RStudio also provides a "Viewer" pane.

Anything interactive will be displayed there.


## Packages

<span class="pack">ggplot2</span> Is the most widely used package for visualization in R.

However, it is not interactive by default.

Many packages use <span class="pack">htmlwidgets</span> and **d3** (javascript library) to provide interactive graphics.


Some packages to note:
  - <span class="pack">plotly</span>
    - used also in Python, Matlab, Julia, aside from many interactive plots, can convert ggplot2 images to interactive ones.

  - <span class="pack">ggvis</span>
    - interactive successort to ggplot though not currently actively developed

  - <span class="pack">rbokeh</span>
    - like plotly, also has cross program support

  - <span class="pack">DT</span>
    - interactive data



## Interactive Tables (View)

RStudio now gives you a sortable table with the <span class="func">View</span> function.

  - <span class="func">View</span> just like you would otherwise use

```{r, eval = FALSE}
View(state.x77)
```

## Example

Works in your presentations too.

```{r}
library(plotly)
df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/2011_february_us_airport_traffic.csv')
df$hover <- with(df, paste(airport, city, state, "Arrivals: ", cnt))

# marker styling
m <- list(
  colorbar = list(title = "Incoming flights February 2011"),
  size = 8, opacity = 0.8, symbol = 'square'
)

# geo styling
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showland = TRUE,
  landcolor = toRGB("gray95"),
  subunitcolor = toRGB("gray85"),
  countrycolor = toRGB("gray85"),
  countrywidth = 0.5,
  subunitwidth = 0.5,
  bgcolor='rgba(0,0,0,0)'
)

plot_ly(df, lat = lat, lon = long, text = hover, color = cnt, marker=m,
        type = 'scattergeo', locationmode = 'USA-states', mode = 'markers', colors='RdBu') %>%
  layout(title = 'Most trafficked US airports<br>(Hover for airport)', geo = g,
         paper_bgcolor='rgba(0,0,0,0)',
         plot_bgcolor='rgba(0,0,0,0)',
         font=list(color=toRGB("gray85")),
         width=1200
         )
```



## Shiny

<span class="emph">Shiny</span> is a framework that can essentially allow you to build an interactive website.

Most of the more recently developed* packges will work specifically within the shiny and rmarkdown settings.


<span style="font-size:10">R has a long history of providing interactive graphics, but most of it was very poor.</span>


## Interactive Tables With 'DT'


The <span class="pack">DT</span> package allows you to create interactive data tables.

These tables will open in your viewer:

  - They also work nicely within R Markdown and similar formats

    - They don't, however, work in ioslides!

## Checking Your Data


It is always a good idea to give your data a visual pass.

```{r, eval = FALSE, echo=-1}
library(dplyr)

DT::datatable(state.x77)
```

Identify anomolies, nonsense values, etc.

## Other Things To Check


Summaries

```{r, eval = FALSE}
cbind(data.frame(state.x77), state.division) %>%
  group_by(state.division) %>%
  summarize(meanPop = round(mean(Population), 2),
            meanMurder = round(mean(Murder), 2)) %>%
  DT::datatable()
```

Correlation Tables

```{r, eval = FALSE}
data.frame(state.x77) %>%
  cor() %>%
  round(2) %>%
  DT::datatable()
```


## Some Additional Functionality


The <span class="func">datatable</span> function also has an argument to filter your data.

```{r, eval = FALSE}
DT::datatable(cbind(data.frame(state.x77), state.division), filter = "top")
```

Same as above, but with division as a string:

```{r, eval = FALSE}
DT::datatable(cbind(data.frame(state.x77), as.character(state.division)), filter = "top")
```

## Interactive Plots


RStudio, in conjunction with some modern packages, lets you visually explore your data.

There are two packages that are great for this:

  - <span class="pack">ggvis</span>

  - <span class="pack">rbokeh</span>

## ggvis

With <span class="pack">ggvis</span>, you can take some of what you know from <span class="pack">ggplot2</span> and use it to create interactive, web-ready plots.

  - Instead of building layers with <span class="func">+</span>, <span class="pack">ggvis</span> imports the <span class="func">%>%</span> operator from <span class="pack">magrittr</span>.


## ggvis


```{r, eval=FALSE}
library(ggvis)

cbind(data.frame(state.x77), state.division) %>%
  ggvis(~ Illiteracy, ~ Income) %>%
  layer_points(fill = ~ state.division) %>%
  layer_smooths(span = input_slider(.2, 1))
```


## rbokeh


The <span class="pack">rbokeh</span> package is a port of the 'bokeh' package from Python.

It is very similar to <span class="pack">ggvis</span>, but there are some distinct differences.

## An Example


```{r, eval=FALSE}
library(rbokeh)

cbind(data.frame(state.x77), state.name) %>%
  figure() %>%
  ly_points(x = Illiteracy, y = Income,
            color = state.division,
            hover = state.name) %>%
  ly_lines(lowess(x = Illiteracy,
                  y = Income))

```


## Which One?


It might be easier to get up and going with <span class="pack">rbokeh</span>.

On the other hand, <span class="pack">ggvis</span> is incredibly powerful and is ready to go into any <span class="pack">shiny</span> app.


## Quick Wrap


RStudio lets you take a deeper look at your data.

Interactive tables and plots go a long way to helping you understand your data better.